Staged typed type-preserving tagless ``interpreters'', \emph{finally}

### 2007-04-06 tagless/tagless-final.tex
Add bibliography entries

We used a method from \citet{Asai-SAS99} to reduce the amount
of back-and-forth between reify/reflect: in our PE we track both
static and dynamic values in parallel (whenever possible).  In some sense,
we are observing that this method type-checks in Hindley-Milner once we
deforest the static code representation.  Put another way, we have
manually applied type-level partial-evaluation to our typelevel 
functions (see \S\ref{S:pe-solution}) to obtain simpler types 
acceptable to MetaOCaml.  It is worthwhile to note that 
\citet{Sumii-HOSC2001} also takes advanatage of this method
\citep[of][]{Asai-SAS99} to get interesting results combining online
and offline partial evaluation.  Like \citet{SperberThiemann:TwoForOne},
we strive for modularity by reusing earlier stages.  It would be interesting
to see if we could derive a \emph{cogen} \citep{Thiemann:cogeninsixlines}
in the same manner.

@@

A lot of effort has gone into finding ways to give precise types to untyped
terms (see for example \citet{baars-typing,Guillemette-Monier-PLPV}
as well as the messages \citet{haskell-019160,haskell-019161}).  There 
take advantage of the host language's types to varying extents.  All 
these approaches are very successful for proper terms, but less so for
ill-typed terms, where error messages can be quite obscure.  We use the
host's typing facilities to do all the work, and thus inherit the strengths
and weaknesses of the host language.  While that may not be optimal, it is
certainly much more modular.

>>>>

We used a method from \citet{asai-binding-time} to reduce the amount
of back-and-forth between reify/reflect: in our PE we track both
static and dynamic values in parallel (whenever possible).  In some sense,
we are observing that this method type-checks in Hindley-Milner once we
deforest the static code representation.  Put another way, we have
manually applied type-level partial-evaluation to our typelevel 
functions (see \S\ref{S:PE-solution}) to obtain simpler types 
acceptable to MetaOCaml.  It is worthwhile to note that 
\citet{sumii-hybrid} also takes advantage of this method
\citep[of][]{asai-binding-time} to get interesting results combining online
and offline partial evaluation.  Like \citet{SperberThiemann:TwoForOne},
we strive for modularity by reusing earlier stages.  It would be interesting
to see if we could derive a \emph{cogen} \citep{Thiemann:cogeninsixlines}
in the same manner.

@@

A lot of effort has gone into finding ways to give precise types to untyped
terms (see for example \citet{baars-typing,Guillemette-Monier-PLPV}
as well as the messages \citet{haskell-list}).  There 
take advantage of the host language's types to varying extents.  All 
these approaches are very successful for proper terms, but less so for
ill-typed terms, where error messages can be quite obscure.  We use the
host's typing facilities to do all the work, and thus inherit the strengths
and weaknesses of the host language.  While that may not be optimal, it is
certainly much more modular.
